The Effects of weather on maize yields: New evidence from Kenya

### 2018-12-17 writing/draft3/draft14.tex
write and so..

\sloppy
Kenya consists of $47$ counties with semi-autonomous county governments  \cite{Barasa2017}. As a result of the high degree of county-level autonomy, the policies and regulations often differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties. Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (also known as mixed models) commonly used to analyse longitudinal data \cite{bates2000mixed}. Mixed models are suitable for analysis of panel data as they account for the panel structure of the dataset. These types of models include both fixed effects parameters and random effects. Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \cite{bates2010lme4}. On the other hand, random effects are unobserved random variables. There are at least three benefits of treating a set of parameters as a random sample from some distribution. \textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation %(\citealp{KERYch9, KERYch12}).

@@

\textcolor{blue}{we checked this approach by drop1 =anova testing and it turns out that our model is better than each model without one of the variables..Also the automatic step procedure, which we run for comparison gave us almost the same model (forcing the seasonal precip and average temperature). The only varibale which was extra was MaxT and we decided to remove it because of high VIF - or correlation with average temp...Also mention add1}

According to a histogram, the shape of our maize yield data was closer to a log-normal than to a normal distribution. Therefore, following the common practice, the  maize yield values were log-transformed before conducting the analysis. Furthermore, we scaled all predictors by subtracting mean and dividing them by standard deviation to avoid convergence problems.

According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$). To further investigate the autocorrelation structure of the errors, we estimated a number of models (with the subset of predictors chosen as described above) with ARMA($p$,$q$) error structure. In particular, we estimated all variants of our model with ARMA($p$,$q$) errors such that $p\leq2$ and $q\leq2$. Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found that the most appropriate error correlation structure was ARMA($1$,$1$). The value of the AIC criterion of the model with ARMA($1$,$1$) error structure was $2122.2$ while the AIC was $2129.2$ for the model with ARMA($1$,$0$) errors.  The p-value of the likelihood ratio test of the comparison of the model with ARMA($1$,$1$) errors and the model with ARMA($1$,$0$) errors was smaller than $1 \times10^{-4}$, hence the ARMA($1$,$1$) error structure turned out to be a better fit. The AIC criteria and the likelihood ratio tests of all the models with ARMA($p$,$q$) errors are summarised in table~\ref{ARMA} in the appendix.

To find the most suitable set of fixed effects, we adapted a stepwise selection procedure to choose a model with the smallest value of the Akaike’s Information Criterion (AIC). We started  with a model which included the complete set of our weather measures in the fixed effects and random intercepts and we performed a search through the subsets off the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (removing and adding predictors). We decided to not to include the number of heatwave days and average maximum daily temperature although they slightly improve the AIC if present. The reason for not including them is that they are strongly correlated with average seasonal temperature (the value of correlation coefficient is $0.981$ for the maximum daily temperature and $0.517$ for the number of heatwave days and the p-values of both of them are smaller than $1\times10^{-6}$) and they are not significant at the level of significance $\alpha=0.05$ (their p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days). Furthermore, including them only improved AIC by $4$ (when estimated by maximum likelihood as required for a meaningful comparison) and the signs of these variables were positive in contrary to our assumption. Hence, we concluded that the decrease in AIC and positive signs are likely to be result of their correlation with average temperature or other omitted variable and we did not include them in our model.

We further applied a series of likelihood ratio tests to test for significance of random slopes, however no subset of random slopes improved the fit, hence, our final model only included intercepts in random effects.

To verify our results, we adapted an alternative step-down model building
approach using the Satterthwaite’s method to determine the p-values of the individual
t-tests \cite{lmerTest}. To be more specific, we started with a model which included the complete set of our weather measures in both fixed effects and random effects. In the first part of this procedure the insignificant random effects were removed one by one until only significant predictors remained. After this, the insignificant fixed effects were removed one by one until only significant predictors remained. For random effects we considered the level of significance $\alpha=0.1$ and for fixed effects we opted for $\alpha=0.05$~\cite{lmerTest}. Using this method we chose the same model as according to the AIC-based stepwise method described above.

>>>>

\sloppy
Kenya consists of $47$ counties with semi-autonomous county governments. As a result of the high degree of county-level autonomy, the policies and regulations often differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties. Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (also known as mixed models) commonly used to analyse longitudinal data \cite{bates2000mixed}. Mixed models are suitable for analysis of panel data as they account for the panel structure of the dataset. These types of models include both fixed effects parameters and random effects. Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \cite{bates2010lme4}. On the other hand, random effects are unobserved random variables. There are at least three benefits of treating a set of parameters as a random sample from some distribution. \textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation %(\citealp{KERYch9, KERYch12}).

@@

According to a shape of a histogram of the maize yield data we decided to log-transform the yield data before conducting our analysis. Hence, we used a log-linear functional form of our models as is a common practice in agricultural economics. Furthermore, we scaled all predictors by subtracting mean and dividing them by standard deviation to avoid convergence problems.

To find the most suitable set of fixed effects, we adapted a stepwise selection procedure to choose a model which minimized the value of the Akaike’s Information Criterion (AIC). We started  with a model which included the complete set of our weather measures in the fixed effects and random intercepts and we performed a search through the subsets off the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (that is removing and adding predictors). We decided to not to include the number of heatwave days and average maximum daily temperature although including them in the fixed effects slightly improved the AIC. The reasons for not including them were their strong correlation with average seasonal temperature (the value of the correlation coefficient was $0.981$ for the correlation of the average seasonal temperature with the maximum daily temperature and $0.517$ the correlation of the average seasonal temperature with the number of heatwave days and the p-values of both of them are smaller than $1\times10^{-6}$) and the lack of significance of their individual t-tests (their p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days). Furthermore, including them only improved AIC by $4$ (when estimated by maximum likelihood as required for a meaningful comparison) and the signs of these variables were positive in contrary to our expectation. Hence, we concluded that the decrease in AIC and positive signs are likely to be results of the correlation with average temperature or perhaps other omitted variables and we did not include them in our model.

We further applied a series of likelihood ratio tests to test for significance of random slopes, however no subset of random slopes improved the fit, hence, our final model only includes random intercepts.

To verify our results, we adapted an alternative step-down model building
approach using the Satterthwaite’s method to determine the p-values of the individual
t-tests \cite{lmerTest}. To be more specific, we started with a model which included the complete set of our weather measures in both fixed effects and random effects. In the first part of this procedure the insignificant random effects were removed one by one until only significant predictors remained. After this, the insignificant fixed effects were removed one by one until only significant predictors remained. For random effects we considered the level of significance $\alpha=0.1$ and for fixed effects we opted for $\alpha=0.05$~\cite{lmerTest}. Applying this method we obtained the same model as according to the AIC-based stepwise method described above.

\textcolor{blue}{re-write the tables with estimates  UNITS KELVIN AND CV TIMES THOUSAND; possibly table of all available weather measure in the appendix, maybe a drop1 and add1 tables??}

Based on the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$). To further investigate the autocorrelation structure of the errors, we estimated a number of models (with the subset of predictors chosen as described above) with ARMA($p$,$q$) error structure. In particular, we estimated all variants of our model with ARMA($p$,$q$) errors such that $p\leq2$ and $q\leq2$. Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found that the most appropriate error correlation structure was ARMA($1$,$1$). The value of the AIC criterion of the model with ARMA($1$,$1$) error structure was $2122.2$ while the AIC was $2129.2$ for the model with ARMA($1$,$0$) errors.  The p-value of the likelihood ratio test of the comparison of the model with ARMA($1$,$1$) errors and the model with ARMA($1$,$0$) errors was smaller than $1 \times10^{-4}$, hence the ARMA($1$,$1$) error structure turned out to be a better fit. The AIC criteria and the likelihood ratio tests of all the models with ARMA($p$,$q$) errors are summarised in table~\ref{ARMA} in the appendix.
